Query Optimization for Dynamic Imputation

### 2017-03-02 paper/algorithm.tex

In general, evaluating the time complexity for different \textit{Impute} operators is
complicated, as the properties of
different algorithms can vary significantly, the underlying I/O cost
is not easily extracted (especially in the case that the entire set of tuples does not
fit in memory), and the CPU cost dominates (in contrast to the other operators, which do
not consider CPU cost explicitly). For example,~\cite{martin1995time} finds that the
time complexity of the \textit{build tree} algorithm for one commonly-used class of
decision trees is a function of the number of classes, several overhead constants, the
parameterization of the partition and heuristic functions, and the \textit{arity} (the
number of subsets considered for each split). The \textit{build tree} phase often
does not even dominate computation, as post-processing steps like pruning, which are vital
in achieving good performance, can have cost exponential in the height of the tree.
Nonetheless, in practice, we find that a simple parameterization can be acceptable (i.e.
yield intelligent query plans).

@@

The plan cache uses the partial order of plans defined by  $\langle \textsc{Dirty}, \textsc{Loss}, \textsc{Time} \rangle$ to
collect sound and complete Pareto frontiers. For plans with the same dirty set, we can treat $\langle \textsc{Loss}, \textsc{Time} \rangle$
as a total order, allowing us to appropriately select the frontier. Meanwhile, plans with different $\textsc{Dirty}$ sets cannot be compared. The final
Pareto frontier produced by the optimizer corresponds to the collection of plans associated with the empty dirty set.

@@

\ProjectName{} is designed so that any imputation strategy
can be plugged in with minimal effort and without changes to the optimizer, simply by adding
time and loss functions (\Cref{sec:cost-model}) which describe the new algorithm.
For our purposes, an imputation strategy is a function $I(R, C) \rightarrow R'$ which fills in the missing
attributes $C$ in a relation $R$, returning a new set of tuples $R'$.
The flexibility of this approach aligns with the spirit of query optimization, which aims to
reconcile performance with a declarative interface to data manipulation.

>>>>

In general, evaluating the time complexity for different \textit{Impute} operators is
complicated, as the properties of
different algorithms can vary significantly, the underlying I/O cost
is not easily extracted (especially in the case that the entire set of tuples does not
fit in memory), and the CPU cost dominates (in contrast to the other operators, which do
not consider CPU cost explicitly). For example,~\cite{martin1995time} finds that the
time complexity of the \textit{build tree} algorithm for one commonly-used class of
decision trees is a function of the number of classes, several overhead constants, the
parameterization of the partition and heuristic functions, and the \textit{arity} (the
number of subsets considered for each split). The \textit{build tree} phase often
does not even dominate computation, as post-processing steps like pruning, which are vital
in achieving good performance, can have cost exponential in the height of the tree.
Nonetheless, in practice, we find that a simple parameterization can be acceptable (i.e.
yields intelligent query plans).

@@

The plan cache uses the partial order of plans defined by  $\langle \textsc{Dirty}, \textsc{Loss}, \textsc{Time} \rangle$ to
collect sound and complete Pareto frontiers. For plans with the same dirty set, we can treat $\langle \textsc{Loss}, \textsc{Time} \rangle$
as a total order, allowing us to appropriately select the frontier. Plans with different $\textsc{Dirty}$ sets cannot be compared. The final
Pareto frontier produced by the optimizer corresponds to the collection of plans associated with the empty dirty set.

@@

\ProjectName{} is designed so that any imputation strategy
can be plugged in with minimal effort and without changes to the optimizer, simply by adding
time and loss functions (\Cref{sec:cost-model}) which describe the new algorithm.
For our purposes, an imputation strategy is a function $I(R, C) \rightarrow R'$ which fills in the missing
attributes $C$ in a relation $R$, returning a new relation $R'$.
The flexibility of this approach aligns with the spirit of query optimization, which aims to
reconcile performance with a declarative interface to data manipulation.
